```{r setup2, include=FALSE}
library(here)
library(dataRetrieval)
library(tidyverse)
library(trend)
library(lubridate)
library(sf)
library(nhdplusTools)
library(mapview)
library(forcats)
library(ggplot2)
library(dplyr)
library(bookdown)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.cap = "")
knitr::opts_chunk$set(show_col_types = FALSE)
```


# Methods

Downloaded turbidity from USGS site number 09014050 that is at the Grand lake outlet that was affected by the East Troublesome Fire at a low burn location.
Downloaded discharge and turbidity from USGS site number 09034250 that is along troublesome creek and outlets into the Colorado river that was affected by the East Troublesome Fire in a high burn area.
Downloaded discharge and turbidity from USGS site number 09040500 that is along the Colorado River that was affected by the East Troublesome Fire in a low burn area.
Downloaded discharge and turbidity from USGS site number 09027100 that is along the Fraser river that wass not affected by East Troublesome Fire in a non burn area


## USGS Sites

```{r}
site_no <- "09034250" 
site_no2 <- "09040500"
site_no3 <- "09014050"

control_site <- "09027100"
```
Site Numbers labeld to help pull data for each considered site.

## Downloaded Discharge Data

```{r}

# Discharge data pulled form the site_no from 10/01/1986 to 1/1/2022
q <- readNWISdv(siteNumbers = site_no,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

q_ind <- q %>%
  mutate(q_in = (((q_cfs / (2.311*10^10))*3600)*12)*24 )

# site_no2
q2 <- readNWISdv(siteNumbers = site_no2,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

q2_ind <- q2 %>%
  mutate(q_in = (((q_cfs / (4.684*10^9))*3600)*12)*24 )

# site_no3
q3 <- readNWISdv(siteNumbers = site_no3,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

q3_ind <- q3 %>%
  mutate(q_in = (((q_cfs / (2.311*10^10))*3600)*12)*24 )

# control site
qC <- readNWISdv(siteNumbers = control_site,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

qC_ind <- qC %>%
  mutate(q_in = (((q_cfs / (3.206*10^9))*3600)*12)*24 )
```
Code used to dowload dischrage from each site in cfs then covnverted in in/d.

## Turbidity 

```{r}
# site 1 daily turbidity 
T1_l <- readNWISuv(siteNumbers = site_no,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))

write_csv(T1_l,'data/T1maddy.csv')

Turbidity_1 <- read_csv('data/T1maddy.csv')%>%
  rename( Date = date)

# compressed dated data
T1_s <- with(T1_l, T1_l[(date >= "2021-06-07" & date <= "2021-10-01"), ])

write_csv(T1_s,'data/T1smaddy.csv')

Turbidity_1s <- read_csv('data/T1smaddy.csv')%>%
  rename( Date = date)

# site 2 daily turbidity 
T2_l <- readNWISuv(siteNumbers = site_no2,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))

write_csv(T2_l,'data/T2maddy.csv')

Turbidity_2 <- read_csv('data/T2maddy.csv')%>%
  rename( Date = date)

# compressed dated data
T2_s <- with(T2_l, T2_l[(date >= "2021-06-07" & date <= "2021-10-01"), ])

write_csv(T2_s,'data/T2smaddy.csv')

Turbidity_2s <- read_csv('data/T2smaddy.csv')%>%
  rename( Date = date)

# site 3 daily turbidity

# longer dated data
T3_l <- readNWISuv(siteNumbers = site_no3,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))


write_csv(T3_l,'data/T3lmaddy.csv')

Turbidity_3l <- read_csv('data/T3lmaddy.csv')%>%
  rename( Date = date)

# compressed dated data
T3_s <- with(T3_l, T3_l[(date >= "2021-06-07" & date <= "2021-10-01"), ])

write_csv(T3_s,'data/T3smaddy.csv')

Turbidity_3s <- read_csv('data/T3smaddy.csv')%>%
  rename( Date = date)


# control site daily turbidity
TC_l <- readNWISuv(siteNumbers = control_site,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))


write_csv(TC_l,'data/TCmaddy.csv')

Turbidity_C <- read_csv('data/TCmaddy.csv')%>%
  rename( Date = date)

# compressed dated data
TC_s <- with(TC_l, TC_l[(date >= "2021-06-07" & date <= "2021-10-01"), ])

write_csv(TC_s,'data/TCsmaddy.csv')

Turbidity_Cs <- read_csv('data/TCsmaddy.csv')%>%
  rename( Date = date)

```
Code downloaded continuous turbidity data, where it then converts the continuous data into daily data. After this code runs I then create a csv file of the daily data and then recall it to my data source. I aslo had to shorten collected dates on certain turbidity levels for combined figure 3.9 to show values during the same time after the fire.  

## Combined Data

### Discharge

```{r}
Q_c <- rbind(q_ind, q2_ind, qC_ind)

Q_cr <-Q_c 

Q_cr$site_no <- fct_recode(Q_cr$site_no ,
                                 "Low Burn " = "09034250",
                                 "High Burn" = "09040500",
                                 "No Burn" = "09027100")

colnames(Q_cr)[1] <- "BurnLevel"

```
Code binds all needed data sets and then re codes the site_no column to rename the site numbers to burn level, and then I renamed the column name to BurnLevel.

### Turbidity

```{r}
T_c <- rbind(Turbidity_1s, Turbidity_2s, Turbidity_3s, Turbidity_Cs)

T_cr <- T_c 

T_cr$site_no <- fct_recode(T_cr$site_no ,
                                 "Low Burn 1" = "09014050",
                                 "Low Burn 2" = "09034250",
                                 "High Burn" = "09040500",
                                 "No Burn" = "09027100")

colnames(T_cr)[1] <- "BurnLevel"

```
Code binds all needed data sets and then re codes the site_no column to rename the site numbers to burn level, and then I renamed the column name to BurnLevel.

## Site Discription

[Low Burn: USGS 09014050 GRAND LAKE OUTLET BLW CHIPMUNK LN AT GRAND LK, CO] [Low Burn: USGS site 09034250 COLORADO RIVER AT WINDY GAP, NEAR GRANBY, CO.] [High Burn: USGS 09040500 TROUBLESOME CREEK NEAR TROUBLESOME, CO.] [Not Burned: USGS 09027100 FRASER RIVER AT TABERNASH, CO]

### Deliniated Watershed
```{r}
watershed <- st_read('Granby.geojson')%>%
  dplyr::filter(id=='globalwatershed')

mapview(watershed)
```
Watershed of the sites 09014050 located at Grand Lake Outlet and 09034250 located at the Colorado River at Windy Gap. Both site are considered a low burn gauge.
```{r}
watershed_2 <- st_read('Troublesome.geojson')%>%
  dplyr::filter(id=='globalwatershed')

mapview(watershed_2)
```
Watershed of site 09040500 located at Troublesome creek. This site is considered a high burn gauge.  
```{r}
watershed_c <- st_read('NewControl.geojson')%>%
  dplyr::filter(id=='globalwatershed')

mapview(watershed_c)
```
Watershed of site 09027100 located at Fraser River. This site is considered a non burned gauge and is the control site.
