```{r setup2, include=FALSE}
library(here)
library(dataRetrieval)
library(tidyverse)
library(trend)
library(lubridate)
library(sf)
library(nhdplusTools)
library(mapview)
library(forcats)
knitr::opts_chunk$set(echo = TRUE)
```


# Methods

Downloaded data from USGS site number 09014050 that is at the Grand lake outlet that was affected by the East Troublesome Fire.
Downloaded data from USGS site number 09034250 that is along troublesome creek and the Colorado river that was affected by the East Troublesome Fire.
Downloaded data from USGS site number 09040500 that is along the Colorado River that was affected by the East Troublesome Fire.


## USGS Sites

```{r}
site_no <- "09034250" 
site_no2 <- "09040500"
site_no3 <- "09014050"

control_site <- "09027100"
```

## Downloaded Discharge Data

```{r}

# Discharge data pulled form the site_no from 10/01/1986 to 1/1/2022
q <- readNWISdv(siteNumbers = site_no,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

q_ind <- q %>%
  mutate(q_in = (((q_cfs / (2.311*10^10))*3600)*12)*24 )

# site_no2
q2 <- readNWISdv(siteNumbers = site_no2,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

q2_ind <- q2 %>%
  mutate(q_in = (((q_cfs / (4.684*10^9))*3600)*12)*24 )

# site_no3
q3 <- readNWISdv(siteNumbers = site_no3,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

q3_ind <- q3 %>%
  mutate(q_in = (((q_cfs / (2.311*10^10))*3600)*12)*24 )

# control site
qC <- readNWISdv(siteNumbers = control_site,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)

qC_ind <- qC %>%
  mutate(q_in = (((q_cfs / (3.206*10^9))*3600)*12)*24 )
```


## Turbidity 

```{r}
# site 1 daily turbidity 
T1 <- readNWISuv(siteNumbers = site_no,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))

write_csv(T1,'data/T1maddy.csv')

Turbidity_1 <- read_csv('data/T1maddy.csv')%>%
  rename( Date = date)

# site 2 daily turbidity 
T2 <- readNWISuv(siteNumbers = site_no2,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))

write_csv(T2,'data/T2maddy.csv')

Turbidity_2 <- read_csv('data/T2maddy.csv')%>%
  rename( Date = date)

# site 3 daily turbidity

# longer dated data
T3_l <- readNWISuv(siteNumbers = site_no3,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))


write_csv(T3_l,'data/T3lmaddy.csv')

Turbidity_3l <- read_csv('data/T3lmaddy.csv')%>%
  rename( Date = date)

# compressed dated data
T3_s <- with(T3_l, T3_l[(date >= "2021-06-07" & date <= "2021-10-01"), ])

write_csv(T3_s,'data/T3smaddy.csv')

Turbidity_3s <- read_csv('data/T3smaddy.csv')%>%
  rename( Date = date)


# control site daily turbidity
TC <- readNWISuv(siteNumbers = control_site,
                parameterCd = '63680')%>%
  mutate(dateTime = lubridate::as_datetime(dateTime),
         date = lubridate::as_date(dateTime))%>%
  group_by(site_no, date)%>%
  summarize(mean = mean(X_63680_00000))


write_csv(TC,'data/TCmaddy.csv')

Turbidity_C <- read_csv('data/TCmaddy.csv')%>%
  rename( Date = date)

```

## Combined Data

### Discharge

```{r}
Q_c <- rbind(q_ind, q2_ind, q3_ind, qC_ind)

Q_cr <-Q_c 

Q_cr$site_no <- fct_recode(Q_cr$site_no ,
                                 "Low Burn 1" = "09014050",
                                 "Low Burn 2" = "09034250",
                                 "High Burn" = "09040500",
                                 "No Burn" = "09027100")

colnames(Q_cr)[1] <- "BurnLevel"

```

### Turbidity

```{r}
T_c <- rbind(Turbidity_1, Turbidity_2, Turbidity_3s, Turbidity_C)

T_cr <- T_c 

T_cr$site_no <- fct_recode(T_cr$site_no ,
                                 "Low Burn 1" = "09014050",
                                 "Low Burn 2" = "09034250",
                                 "High Burn" = "09040500",
                                 "No Burn" = "09027100")

colnames(T_cr)[1] <- "BurnLevel"
```

## Site Discription

USGS 09014050 GRAND LAKE OUTLET BLW CHIPMUNK LN AT GRAND LK, CO
USGS site 09034250 COLORADO RIVER AT WINDY GAP, NEAR GRANBY, CO.
USGS 09040500 TROUBLESOME CREEK NEAR TROUBLESOME, CO.

### Deliniated Watershed
```{r}
watershed <- st_read('Granby.geojson')%>%
  dplyr::filter(id=='globalwatershed')

mapview(watershed)

watershed_2 <- st_read('Troublesome.geojson')%>%
  dplyr::filter(id=='globalwatershed')

mapview(watershed_2)

watershed_c <- st_read('NewControl.geojson')%>%
  dplyr::filter(id=='globalwatershed')

mapview(watershed_c)
```

