---
title: "Wildfires Affect on East Troublesome Fire"
author: "Madison Thompson"
date: "3/3/2022"
site: bookdown::bookdown_site
documentclass: docs
output:
  bookdown::gitbook: default
bibliography: citations.bib
---

# Introduction 

```{r setup, include=FALSE}
library(here)

knitr::opts_chunk$set(echo = TRUE)
```

 Wildfires put a long term scare on the land that has many major effects, one being
  the state of the watershed the fire takes place in. Wildfires were once a common natural
  occurrence in most forest ecosystems that did not have a devastating effect on the land.
  But in more recent years things such as climate change, fuel load accumulation, extensive 
  droughts, and an increase in human presence in forests has caused a devastating effect on 
  the land when these wildfires burn. The concern that comes from these factors has put a heavier
  strain on the understanding hoe the landscape changes and reacts to the wildfire. In particular, 
  there is a need to understand the interaction between wildfire, and associated responses in 
  vegetation cover, and hydrologic and geomorphic behavior. [@owens_muted_2013]
  
  When a wildfire burns it way through the land it often causes dramatic changes in forest vegetation 
  structure and soil conditions, it also alters the watershed processes that control streamflow, 
  soil erosion, nutrient export, and downstream water chemistry [@hohner_wildfires_2019].The effect
  of erosion, infiltration, runoff, and slope failure play a major role in how the land is re-molded. 
  Where this re-modeling creates an everlasting change to the watershed in the area. These changes can 
  include the streamflow, landscape, and the soil of the land. High-severity fires heat the mineral soil 
  and consume all vegetative ground cover, leaving only ash at the soil surface [@kunze_streamflow_2006].
  When the fire creates these changes to the land it in turn effects the watershed and flow of water. These 
  effects that occur create a domino effect that starts with the forest ecosystem and begins to alter
  the land and watershed from the most obvious to the least observed occurrences. 

  After the wildfires burn and cause the changes to the watersheds soil and hydrologic system 
  it then leads to further effects on the land. Sediment load, flooding,runoff and erosion create
  new factors to consider. These large erosion events, such as those caused by wildfire, are 
  short-lived on geological time scales, but can dominate the long-term sediment yield 
  response[@murphy_effects_nodate]. This is an important factor to consider due to the effect 
  sediment load has on water quality. The mineralization of organic matter, interruption of 
  root uptake, and loss of shade can further impact water quality by increasing stream 
  temperatures and nutrient concentrations [@ice_effects_nodate]. The postfire effects that 
  is seen on the water quality has to do with water surface composition and the drinking water 
  treatment. Once down-stream the water quality effects is better seen at water treatment plants 
  and what needs to be done now to purify the water that was affected by postfire conditions. 
  Extreme postfire erosion can damage infrastructure, accumulate in reservoirs, and challenge 
  coagulation, filtration, and solids handling processes [@hohner_wildfires_2019]. With all these
  factors occurring it causes these water treatment facilities to adapt to these new quality
  features so that the people dependent on that water source can still use the water source. 
  Changes in the quantity and composition of dissolved organic matter (DOM) exported from burned 
  watersheds can influence treatment process performance and the formation of disinfection byproducts 
  (DBPs) [@hohner_wildfires_2019]. Here it is seen that even after the fire is out it does not mean
  the effects of the fire is gone. The wildfire goes on to create issues within the water quality
  of the watershed.
  
  In the state of Colorado warmer spring air temperatures and longer wildfire seasons 
  in the western US [@mast_water-quality_2016] had created the land for larger and
  longer wildfire season. This increase has created concern for larger fires to appear 
  across Colorado's Front Range. With the possibility of even larger fires to occur it creates
  instability in the land once the fire burn through it, which creates alterations to the
  watersheds associated with the fire. Postfire the largest change in the land is the increase 
  in runoff and erosion. The effects that come from runoff and erosion has pushed for efforts to better
  understand wildfires in the Colorado Front Range and effects it has on the soil and water. An example of 
  a wildfire that has affected a watershed was that of the East Troublesome Fire that occurred the afternoon
  of October 14, 2020. Fire behavior played a large role in the severity of impacts to soil in the East 
  Troublesome fire [@us_national_forest_east_2020]. The fire also altered the watershed of this
  area due to the location of this wildfire. The fire burned at a high elevation which is composed
  of a dominated snowmelt watershed attached to the Upper Colorado River. The watershed response 
  within the burned area will likely include an initial flush of ash, rill and gully erosion in 
  headwater drainages and on steep slopes, areas with snowmelt peak flows elevated over 50% relative 
  to pre-fire, sediment-laden flash floods following high-intensity rain events, and potentially 
  debris flows [@us_national_forest_east_2020]. This effect on the watershed can affect 
  the soil, land, and water quality. The streamflow is also highly affected by such a fire due to
  the debris that fall into the streams. Debris flows are among the most hazardous consequences of
  rainfall on burned hillslopes [@us_national_forest_east_2020]. These kinds of debris can affect
  the streamflow and effect the watershed associated with the effected areas. This study will be looking more 
  into the East Troublesome Fire and the effects it has had on the soil and water quality of the watershed.
  

<!--chapter:end:index.Rmd-->

```{r setup, include=FALSE}
library(here)
library(dataRetrieval)
library(tidyverse)
library(trend)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```


# Methods

Downloaded data from USGS site along the Colorado River that was affected by the East Troublesome Fire.

## Downloaded data

```{r}
site_no <- "09034250" 

q <- readNWISdv(siteNumbers = site_no,
                parameterCd = '00060',
                startDate = '1986-10-01',
                endDate = '2022-01-01') %>%
  rename(q_cfs = X_00060_00003) %>%
  addWaterYear(.) %>%
  select(-agency_cd, -X_00060_00003_cd)
```

### Mean monthly discharge 
```{r}
q_month_means <- q%>%
  mutate(month = month(Date),
         year = year(Date))%>%
  group_by(year, month)%>%
  summarize(q_mean = mean(q_cfs, na.rm=T))

head(q_month_means)
```

### Min monthly discharge
```{r}
q_month_mins <- q%>%
  mutate(month = month(Date),
         year = year(Date))%>%
  group_by(year, month)%>%
  summarize(q_min = min(q_cfs))

head(q_month_mins)
```

### Median monthly discharge
```{r}
q_month_medians <- q%>%
  mutate(month = month(Date),
         year = year(Date))%>%
  group_by(year, month)%>%
  summarize(q_median = median(q_cfs))

head(q_month_medians)
```

<!--chapter:end:1_methods.Rmd-->



`r if (knitr::is_html_output()) '# References {-}'`

<!--chapter:end:wrapping_up.Rmd-->

