
# Results

## Discharge Data

```{r, fig.cap = 'Discharge (in/day) recorded from 1986-10-01 to 2022-01-01. Site is low burned and located near Grandby, CO.' }

# Plot for discharge 
{ q1 = "Discharge (in/day) recorded from 1986-10-01 to 2022-01-01. Site is low burned and located near Grandby, CO."}

ggplot(q_ind, aes(x = Date, y = q_in))+
  geom_line()+
  labs( title = 'Discharge at Colorado River Gauge',
        subtitle = 'USGS site 09034350',
        x = 'Year',
        y = 'Q (in/day)' )
```
Figure 3.1 represents the Discharge (Q) levels measured in in/day at the USGS site 09034350 located at the Colorado River at Windy Gap, near Grandby, CO. Due to the location of the fire this site is considered to be in a low burned area. As seen in the figure Q was at a higher level just before the fire in 2010 to 2011, Then around the time of the fire in 2020 and the year after the discharge levels became very low   
```{r, fig.cap='Discharge (in/day) recorded from 2021-05-27 to 2021-12-09. Site is a high burn and is located near Grandby, CO.'}
# site_no2
ggplot(q2_ind, aes(x = Date, y = q_in))+
  geom_line()+
  labs( title = 'Discharge at Troublesome Creek Gauge',
        subtitle = 'USGS site 09040500',
        x = 'Year',
        y = 'Q (in/day)')
```
Figure 3.2 represents Discharge (Q) levels measured at the USGS site 09040500 located at Troublesome Creek,near Grandby, CO. Due to the location of of the fire this site is considered a high burn area. Data was only collected The data collected here was for only on year after the fire occurred and showed overall low discharge values. 

```{r, fig.cap='Discharge (in/day) recorded from 2011-05-01 to 2022-01-01. Site is in a non burned area and is the control site.'}
# control site
ggplot(qC_ind, aes(x = Date, y = q_in))+
  geom_line()+
  labs( title = 'Discharge at Fraser River Gauge',
        subtitle = 'USGS site 09027100',
        x = 'Year',
        y = 'Q (in/day)')
```
Figure 3.3 represents Discharge (Q) levels measured at the USGS site 09027100 located at Fraser River near Tabernash, CO. This site was chosen as the control site since it was located away from the fire so the discharge rate were more consistent and had higher discharge rates  

## Turbidity

```{r, fig.cap='Turbidity data from USGS site 09034250 COLORADO RIVER AT WINDY GAP, NEAR GRANBY, CO. Reccorded time 2021-06-22 to 2021-10-01. USGS gauge located in low burn area'}
# site 1
ggplot(Turbidity_1s, aes(x = Date, y = mean)) + 
  geom_line(size = 1) + 
  labs(title = 'Turbidity Levels Colorado River',
       subtitle = 'USGS site 09034250',
       x = 'Date',
       y = 'T (formazin nephelometric units (FNU))')
```
Figure 3.4 represents the Turbidity levels measured in formazin nephelometric units (FNU) at the USGS site 09034350 located at the Colorado River at Windy Gap, near Grandby, CO. Due to the location of the fire this site is considered to be in a low burned area. The data was collected the year after the fire occurred and showed a consistent growing turbidity rates from the month of June to October, where it peaked between August to September and September to October.  

```{r, fig.cap='Turbidity data from USGS site 09040500 TROUBLESOME CREEK NEAR TROUBLESOME, CO. Reccorded time 2021-06-07 to 2021-10-01. USGS gauge located in high burn area.'}
# site 2
ggplot(Turbidity_2s, aes(x = Date, y = mean)) + 
  geom_line(size = 1) + 
  labs(title = 'Turbidity Levels Troublesom Creek',
       subtitle = 'USGS site 09040500',
       x = 'Date',
       y = 'T (formazin nephelometric units (FNU))')
```
Figure 3.5 represents Turbidity levels measured in formazin nephelometric units (FNU) measured at the USGS site 09040500 located at Troublesome Creek,near Grandby, CO. Due to the location of of the fire this site is considered a high burn area. The data was collected the year after the fire occurred and showed a consistent growing turbidity rates from the month of June to October. This figure presents the highest turbidity values of all three four recorded sites.

```{r, fig.cap='Turbidity data from USGS site 09014050 GRAND LAKE OUTLET BLW CHIPMUNK LN AT GRAND LK, CO. Recorded time from 2010-04-13 to 2021-10-01. USGS gauge located in low burn area.'}
# site 3
ggplot(Turbidity_3l, aes(x = Date, y = mean)) + 
  geom_line(size = 0.5) + 
  labs(title = 'Turbidity Levels Grand Lake Outlet',
       subtitle = 'USGS site 09014050',
       x = 'Date',
       y = 'T (formazin nephelometric units (FNU))')
```
Figure 3.6 represents  Turbidity levels measured in formazin nephelometric units (FNU) measured at the USGS site 09014050 located at the Grand Lake Outlet near Grandby, CO. Due to the location of this site it is considered to be in a low burn area. This figure shows years of turbidity levels in the lake from the year 2010 to 2021.

```{r, fig.cap='Turbidity data from USGS site 09060799 FRASER RIVER AT TABERNASH, CO. Recorded time from 2021-07-15 to 2022-04-07. USGS gauge loctaed in no burn (contro site)'}
#control site
ggplot(Turbidity_Cs, aes(x = Date, y = mean)) + 
  geom_line(size = 1) + 
  labs(title = 'Turbidity Levels Fraser River',
       subtitle = 'USGS site 09070500',
       x = 'Date',
       y = 'T (formazin nephelometric units (FNU))')
```
Figure 3.7 represents  Turbidity levels measured in formazin nephelometric units (FNU) measured at the USGS site 09027100 located at Fraser River near Tabernash, CO. This site was chosen as the control site since it was located away from the fire and had constant lower turbidity levels with only one major spike in August.

## Combined Data

#### Discharge

```{r, fig.cap='Discharge (in/day) from Troublesome Creek (high burn), Colorado River (low burn), and Fraser River (no burn)'}
ggplot(Q_cr, aes(x= BurnLevel, y= q_in))+
  geom_boxplot()+
  scale_y_log10()+
  labs( title = 'Discharge of Sites',
        x = ' Site Burn Level',
        y = 'Q (in/day)')

```
Figure 3.8 represents a combined showing of all three sites discharge levels in box plot form to better show the distribution between the three sites. As seen in the figure discharge values decrease as the severity of the burned fire are increases. 

### Turbidity

```{r, fig.cap='Turbidity levels from Troublesome Creek (high burn), Colorado River (low burn 2), Grand lake Outlet (low burn 1) and Fraser River (no burn'}

ggplot(T_cr, aes(x= Date, y= mean, color = BurnLevel ))+
  geom_point()+
  geom_line()+
  scale_y_log10()+
  labs(title = "Turbidity Levels at Four Sites",
       x = "Month",
       y = "Turbidity (FNU)")
```
Figure 3.9 represents the combined turbidity values from June to October in the year 2021. As seen in the graph high burn severity has a high influence in increased turbidity, while low burn sites are not as greatly infulenced when comparing to the control site variable. 

